---
title: "CS300 Proj2"
author: "Diyang Yu"
date: "April 1, 2017"
output:
  word_document: default
  html_document:
    highlight: tango
    theme: cerulean
header-includes:
- \usepackage{bbm}
- \usepackage{amsmath}
---

```{r,warning=FALSE,message=FALSE}
# Get the pixels values. Now you will have 50 x 50 pixels.

library(jpeg)
jpg1 <- readJPEG("face1.jpg")
jpg2 <- readJPEG("face2.jpg")
jpg3 <- readJPEG("face3.jpg")
jpg4 <- readJPEG("face4.jpg")
jpg5 <- readJPEG("face5.jpg")
jpg6 <- readJPEG("face6.jpg")
jpg7 <- readJPEG("face7.jpg")
jpg8 <- readJPEG("face8.jpg")
jpg9 <- readJPEG("face9.jpg")
jpg10 <- readJPEG("face10.jpg")

```

# Orginal pictures

```{r}
grays = rgb(red = 0:255/255, blue = 0:255/255, green = 0:255/255)
heatmap(jpg1,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg2,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg3,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg4,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg5,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg6,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg7,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg8,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg9,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(jpg10,Rowv=NA,Colv=NA,col=grays, scale = "none")

# Represent every image I as a vector T. Now for each image you will have n^2x1 vector.

vec1 <- as.vector(jpg1)
vec2 <- as.vector(jpg2)
vec3 <- as.vector(jpg3)
vec4 <- as.vector(jpg4)
vec5 <- as.vector(jpg5)
vec6 <- as.vector(jpg6)
vec7 <- as.vector(jpg7)
vec8 <- as.vector(jpg8)
vec9 <- as.vector(jpg9)
vec10 <- as.vector(jpg10)

#Compute the average face vector and subtract the mean face.

bigvec <- cbind(vec1,vec2,vec3,vec4,vec5,vec6,vec7,vec8,vec9,vec10)
mean <- rowMeans(bigvec)

newvec1 <- vec1 - mean
newvec2 <- vec2 - mean
newvec3 <- vec3 - mean
newvec4 <- vec4 - mean
newvec5 <- vec5 - mean
newvec6 <- vec6 - mean
newvec7 <- vec7 - mean
newvec8 <- vec8 - mean
newvec9 <- vec9 - mean
newvec10 <- vec10 - mean
```

# Display the pictures.

```{r}
dim(newvec1) <- c(50,50)
dim(newvec2) <- c(50,50)
dim(newvec3) <- c(50,50)
dim(newvec4) <- c(50,50)
dim(newvec5) <- c(50,50)
dim(newvec6) <- c(50,50)
dim(newvec7) <- c(50,50)
dim(newvec8) <- c(50,50)
dim(newvec9) <- c(50,50)
dim(newvec10) <- c(50,50)
dim(mean) <- c(50,50)

heatmap(newvec1,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec2,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec3,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec4,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec5,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec6,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec7,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec8,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec9,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(newvec10,Rowv=NA,Colv=NA,col=grays, scale = "none")
heatmap(mean,Rowv=NA,Colv=NA,col=grays, scale = "none")
```

```{r}
newmatrix <- cbind(as.vector(newvec1),as.vector(newvec2),as.vector(newvec3),
                   as.vector(newvec4),as.vector(newvec5),as.vector(newvec6),
                   as.vector(newvec7),as.vector(newvec8),as.vector(newvec9),as.vector(newvec10))
```

# Calculate covariance and eigen values.

```{r}
covariance <- cov(newmatrix)
eigenv <- eigen(covariance)
eigenvalue <- eigenv $ values
eigenvector <- eigenv $vectors
plot(eigenvalue,type="l")
```

# Calculate the eigenvectors and display each face with respect to eigenvectors.

```{r}
eigenvalue1 <- eigenvalue[1]
eigenvalue2 <- eigenvalue[2]
eigenvalue3 <- eigenvalue[3]
eigenface1 <- eigenvalue1 * newmatrix
eigenface2 <- eigenvalue2 * newmatrix
eigenface3 <- eigenvalue3 * newmatrix
eigenface <- eigenface1 + eigenface2 + eigenface3

## eigenface 1
eigenpic1 <- matrix(ncol=50, nrow=500)
a <- matrix(ncol=1, nrow=2500)

for (i in 1:10){
  a<- eigenface1[,i]
  dim(a)<-c(50,50)
  eigenpic1[(50*(i-1)+1):(50*i),] <- a
}

for (i in 1:10){
  heatmap(eigenpic1[(50*(i-1)+1):(50*i),],Rowv=NA,Colv=NA,col=grays, scale = "none")
}

## eigenface 2
eigenpic2 <- matrix(ncol=50, nrow=500)
a <- matrix(ncol=1, nrow=2500)

for (i in 1:10){
  a<- eigenface2[,i]
  dim(a)<-c(50,50)
  eigenpic2[(50*(i-1)+1):(50*i),] <- a
}

for (i in 1:10){
  heatmap(eigenpic2[(50*(i-1)+1):(50*i),],Rowv=NA,Colv=NA,col=grays, scale = "none")
}


## eigenface 3
eigenpic3 <- matrix(ncol=50, nrow=500)
a <- matrix(ncol=1, nrow=2500)

for (i in 1:10){
  a<- eigenface3[,i]
  dim(a)<-c(50,50)
  eigenpic3[(50*(i-1)+1):(50*i),] <- a
}

for (i in 1:10){
  heatmap(eigenpic3[(50*(i-1)+1):(50*i),],Rowv=NA,Colv=NA,col=grays, scale = "none")
}


## eigenface
eigenpic <- matrix(ncol=50, nrow=500)
a <- matrix(ncol=1, nrow=2500)

for (i in 1:10){
  a<- eigenface[,i]
  dim(a)<-c(50,50)
  eigenpic[(50*(i-1)+1):(50*i),] <- a
}


for (i in 1:10){
  heatmap(eigenpic[(50*(i-1)+1):(50*i),],Rowv=NA,Colv=NA,col=grays, scale = "none")
}

```

Since there is no much difference in eigen values, there is no much difference in the eigen faces. Adding these values together does not improve the performance a lot. 

# Test
## First test
First, test a picture that is exactly in the training.

```{r}
test1 <- readJPEG("face2.jpg")
featuretest1 <- as.vector(test1) - as.vector(mean)
dis1 <- matrix(nrow=2500, ncol=10)
distance1 <- c()
for (i in 1:10) {
  dis1[,i] <- eigenvalue1 * featuretest1 -eigenface1[,i]
  distance1[i] <- sum(abs(dis1[,i]))/2500
}

distance1
which.min(distance1)
distance1[which.min(distance1)]
```

This exactly meets the original picture. The distance is 0. 

```{r}
dis1 <- matrix(nrow=2500, ncol=10)
distance1 <- c()
for (i in 1:10) {
  dis1[,i] <- eigenvalue2 * featuretest1 -eigenface2[,i]
  distance1[i] <- sum(abs(dis1[,i]))/2500
}

distance1
which.min(distance1)
distance1[which.min(distance1)]
```

When we use the second best engin value, the result is still the same, and the distance is still 0.

## Second test

```{r}
test2 <- readJPEG("face12.jpg")
test2 <- test2[,,1]
heatmap(test2,Rowv=NA,Colv=NA,col=grays, scale = "none")


featuretest2 <- as.vector(test2) - as.vector(mean)
dis2 <- matrix(nrow=2500, ncol=10)
distance2 <- c()
for (i in 1:10) {
  dis2[,i] <- eigenvalue1 * featuretest2 -eigenface1[,i]
  distance2[i] <- sum(abs(dis2[,i]))/2500
}

distance2
which.min(distance2)
distance2[which.min(distance2)]
```


```{r}
dis2 <- matrix(nrow=2500, ncol=10)
distance2 <- c()
for (i in 1:10) {
  dis2[,i] <- eigenvalue2 * featuretest2 -eigenface2[,i]
  distance2[i] <- sum(abs(dis2[,i]))/2500
}

distance2
which.min(distance2)
distance2[which.min(distance2)]
```

When I plug in another my picture from a direction, it is recognized as the forth person, but the differneve between each picture is really small, the distance difference between 2nd and 4th is 0.04.

## Third test

```{r}
test3 <- readJPEG("face11.jpg")
heatmap(test3,Rowv=NA,Colv=NA,col=grays, scale = "none")


featuretest3 <- as.vector(test3) - as.vector(mean)
dis3 <- matrix(nrow=2500, ncol=10)
distance3 <- c()
for (i in 1:10) {
  dis3[,i] <- eigenvalue1 * featuretest3 -eigenface1[,i]
  distance3[i] <- sum(abs(dis3[,i]))/2500
}

distance3
which.min(distance3)
distance3[which.min(distance3)]
```


